---
title: "Wine Quality Manual Model Building"
output: html_notebook
---
## Brief

Investigate the wine_quality_red and wine_quality_white datasets, and, in particular, to model the quality of the wines. Use regression to determine which physiochemical properties make a wine ‘good’. 

Could use leaps or glmulti but doing a manual model

Cleaning below:

```{r}
library(tidyverse)
library(GGally)
library(modelr)
library(janitor)
library(ggfortify)
```

```{r}
wine_red <- read_csv("data/wine_quality_red.csv")
wine_white <- read_csv("data/wine_quality_white.csv")
head(wine_red)
head(wine_white)
```
Both are in the same format so adding a wine_colour col. Could look at each seperately but deciding to combine.

```{r}
wine_red <- wine_red %>% 
  mutate(wine_colour = 'red', .before = fixed_acidity)

wine_white <- wine_white %>% 
  mutate(wine_colour = 'white', .before = fixed_acidity)
```


```{r}
# Extracting the final wine_id of wine_red
last_red_row <- nrow(wine_red)

# Mutating the wine_id's of wine_white by adding this value to each
wine_white <- wine_white %>% 
  mutate(wine_id = wine_id + last_red_row)

# Binding so each wine has a unique id in the final dataframe
wine_both <- bind_rows(wine_red, wine_white)
#View(wine_both)

# Removing wine_id
wine_both <- wine_both %>% 
  select(-wine_id)
```

Checking how many different levels of the categorical variable region there are.

```{r}
wine_both %>%
  distinct(region) %>%
  summarise(number_of_regions = n())
```

```{r}
wine_both <- wine_both %>% 
  mutate(region = as.factor(region),
         wine_colour = as.factor(wine_colour))

class(wine_both$region)
class(wine_both$wine_colour)
```

Checking for aliased variables:

```{r}
alias(quality ~ ., data = wine_both)
```
Variable selection:

```{r}
wine_both %>%
  select(quality, fixed_acidity, volatile_acidity, citric_acid, p_h) %>%
  ggpairs() + 
   theme_grey(base_size = 8) 
```
Most highly correlated with quality is volatile_acidity at -0.25. 

```{r}
wine_both %>%
  select(quality, residual_sugar, chlorides, free_sulfur_dioxide, total_sulfur_dioxide, sulphates, density, alcohol) %>%
  ggpairs() +
  theme_grey(base_size = 1) + 
  theme(strip.text = element_text(size = 5)) 
```

Alcohol and density have a pretty strong positive correlation of almost 0.7 so will drop density from the model.

total_sulfur_dioxide is highly correlated with free_sulfur_dioxide so will drop one.

```{r}
wine_both <- wine_both %>%
  select(-density, -free_sulfur_dioxide)
```

Looking at categorical variables

```{r}
wine_both %>%
  select(quality, wine_colour, region) %>%
  ggpairs() + 
   theme_grey(base_size = 8)
```

alcohol, chlorides, and volatile_acidity seem reasonable

```{r}
model1a <- lm(quality ~ alcohol, data = wine_both)

autoplot(model1a)
```
```{r}
summary(model1a)
```

```{r}
model1b <- lm(quality ~ chlorides, data = wine_both)

autoplot(model1b)

summary(model1b)
```
Not as good

```{r}
model1c <- lm(quality ~ total_sulfur_dioxide, data = wine_both)

autoplot(model1c)

summary(model1c)
```

Re-running ggpairs() with the residuals for alcohol model.

```{r}
wine_remaining_resid <- wine_both %>%
  add_residuals(model1a) %>%
  select(-c("quality", "alcohol"))

ggpairs(wine_remaining_resid) + 
  theme_grey(base_size = 1) +
  theme(strip.text = element_text(size = 5)) 
```

Looks like volatile_acidity and residual_sugar could show something.

```{r}
model2a <- lm(quality ~ alcohol + chlorides, data = wine_both)

autoplot(model2a)
summary(model2a)
```

```{r}
model2b <- lm(quality ~ alcohol + volatile_acidity, data = wine_both)

autoplot(model2b)
summary(model2b)
```
```{r}
model2c <- lm(quality ~ alcohol + residual_sugar, data = wine_both)

autoplot(model2c)
summary(model2c)
```
model2b performs best, using ANOVA to see if addition is justifiable 

```{r}
anova(model1a, model2b)
```
keeping it and moving ot next variable.

```{r}
wine_remaining_resid <- wine_both %>%
  add_residuals(model2b) %>%
  select(-c("quality", "alcohol", "volatile_acidity"))

ggpairs(wine_remaining_resid) + 
   theme_grey(base_size = 1) + 
   theme(strip.text = element_text(size = 5)) 
```

contender variables look to be sulphates and residual_sugar

```{r}
model3a <- lm(quality ~ alcohol + volatile_acidity + sulphates, data = wine_both)

autoplot(model3a)
summary(model3a)
```
```{r}
model3b <- lm(quality ~ alcohol + volatile_acidity + residual_sugar, data = wine_both)

autoplot(model3b)
summary(model3b)
```
```{r}
# Justified in adding this third predictor
anova(model2b, model3a)
```
Next variable:

```{r}
wine_remaining_resid <- wine_both %>%
  add_residuals(model3a) %>%
  select(-c("quality", "alcohol", "volatile_acidity", "sulphates"))

ggpairs(wine_remaining_resid) + 
   theme_grey(base_size = 1) + 
   theme(strip.text = element_text(size = 5))
```

residual_sugar is the only likely addition 

```{r}
model4 <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar, data = wine_both)

autoplot(model4)
summary(model4)
```
```{r}
# Justified in adding this fourth predictor
anova(model3a, model4)
```
Leaving at 4 predictors as to not impact the overall R^2 further.

Moving on to pair interactions of which there are 6:


    alcohol:volatile_acidity
    alcohol:sulphates
    alcohol:residual_sugar
    volatile_acidity:sulphates
    volatile_acidity:residual_sugar
    sulphates:residual_sugar

```{r}
model5pa <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + alcohol:volatile_acidity, data = wine_both)
summary(model5pa)
```

```{r}
model5pb <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + alcohol:sulphates, data = wine_both)
summary(model5pb)
```

```{r}
model5pc <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + alcohol:residual_sugar, data = wine_both)
summary(model5pc)
```

```{r}
model5pd <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + volatile_acidity:sulphates, data = wine_both)
summary(model5pd)
```

```{r}
model5pe <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + volatile_acidity:residual_sugar, data = wine_both)
summary(model5pe)
```

```{r}
model5pf <- lm(quality ~ alcohol + volatile_acidity + sulphates + residual_sugar + sulphates:residual_sugar, data = wine_both)
summary(model5pf)
```

model5pe with the alcohol, volatile_acidity, sulphates, residual_sugar, and volatile_acidity:residual_sugar is the best

```{r}
anova(model4, model5pe)
```
final model is:

quality ~ alcohol + volatile_acidity + sulphates + residual_sugar +   volatile_acidity:residual_sugar

