---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(here)

simd <- read.csv("../clean_data/simd_clean.csv")

```
SIMD  by Admission type


Emergency admissions drive the volume - simd 1 and 2 tend to be admitted more often that otehr SIMD groups - across Scotland
```{r}
library(ggh4x) 
 # and then this updated according to your data set 

simd<- simd %>% drop_na(SIMD)
simd %>%  
  #& HBName == "NHS Ayrshire and Arran") %>% 
group_by(Year, Quarter_single, AdmissionType, SIMD) %>% summarise(number_admissions = sum(Stays)) %>% 
  ggplot(aes(x = interaction(Quarter_single, Year), y = number_admissions, group = SIMD, colour = SIMD))+
   scale_x_discrete(guide = "axis_nested")+
   facet_wrap( ~ AdmissionType)+
  geom_vline(xintercept = "2020Q2", size = 1.5, colour = "red")+
  annotate("rect", xmin = 1, xmax = 3, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue" ) +
  annotate("rect", xmin = 5, xmax = 7, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue") +
  annotate("rect", xmin = 9, xmax = 11, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue") +
  annotate("rect", xmin = 13, xmax = 15, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue") +
  annotate("rect", xmin = 17, xmax = 19, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue") +
  annotate("rect", xmin = 21, xmax = 22, ymin = 0, ymax = 600, alpha = 0.2, fill = "lightblue") +
  geom_line(size = 1.5)+
  theme_light()+
  scale_colour_manual(values = c("#1F3F49","#488A99" ,"#1C4E80","#4cb5f5" ,"#A5d8DD"  ))+
theme(axis.text.x = element_text(hjust = 1, size = 12),
      axis.text.y = element_text(size = 20),
      title  = element_text(size = 20, face = "bold"), 
 
      )+
 
     labs(
    title = "\nHospital Admissions by SIMD\n",
    x = "\nQuarter\n",
    y = "\nNumber of Admissions\n"
  )
  
```


hospital admissions per Health Board per SIMD
```{r}

simd %>% filter(HBName == "NHS Ayrshire and Arran" & SIMD == 4) %>% 
  ggplot(aes(x = Quarter_single, y = AverageLengthOfStay, fill = AdmissionType, colour = AdmissionType)) +
  
  geom_col(position = "dodge", colour = "white")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    facet_wrap(~ Year)+
    labs(
    title = "\nAverage length of hospital stay by SIMD\n",
    x = "\nQuarter\n",
    y = "\nAverage length of stay (days)\n")
    
```

```{r}
# filtered by HBName and SIMD level
simd %>% filter(HBName == "NHS Borders" & SIMD == 1) %>% 
  ggplot(aes(x = Quarter, y = AverageLengthOfStay, group = AdmissionType, colour = AdmissionType)) +
  geom_vline(xintercept = "2020Q2", size = 1.5, colour = "red")+
    annotate("rect", xmin = 1, xmax = 3, ymin = 0, ymax =10, alpha = 0.2, fill = "lightblue" ) +
    annotate("rect", xmin = 5, xmax = 7, ymin = 0, ymax = 10, alpha = 0.2, fill = "lightblue") +
    annotate("rect", xmin = 9, xmax = 11, ymin = 0, ymax = 10, alpha = 0.2, fill = "lightblue") +
    annotate("rect", xmin = 13, xmax = 15, ymin = 0, ymax = 10, alpha = 0.2, fill = "lightblue") +
    annotate("rect", xmin = 17, xmax = 19, ymin = 0, ymax = 10, alpha = 0.2, fill = "lightblue") +
    annotate("rect", xmin = 21, xmax = 22, ymin = 0, ymax = 10, alpha = 0.2, fill = "lightblue") +
  geom_point()+
  geom_line(size = 1)+
  theme_minimal()+
    scale_colour_manual(values = c("#1F3F49","#488A99" ,"#1C4E80","#4cb5f5" ,"#A5d8DD"  ))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
      axis.text.y = element_text(size = 16),
      title  = element_text(size = 14, face = "bold")
      )+
     labs(
    title = "\nHospital Admissions by Health Board, Admission Type and SIMD\n",
    x = "\nQuarter\n",
    y = "\nAverage length of stay (days)\n"
  )
  
  
```
emergency admissions in SIMD 1 and Simd 2 - length of stay

```{r}
simd %>% (distinct(SIMD)
```

SIMD - pre-covid and post-covid emergency admissions have always been at the higher level for SIMD 1 and 2  than SIMD 4 or 5

planned admissions were at similar levels regardless of simd and pre-, after-covid

emergency appoitments drive the pressures on NHS across the entire year


```{r}

simd <- simd %>% filter(Quarter_single == "Q1" | Quarter_single == "Q4") %>% 
  mutate(Covid = case_when(
  Quarter %in% c("2017Q4","2018Q1","2018Q2","2018Q3", "2018Q4", "2019Q1", "2019Q2", "2019Q3", "2019Q4", "2020Q1") ~ "Pre-Covid 19",
    Quarter %in% c("2020Q2","2020Q3","2020Q4","2021Q1","2021Q2","2021Q3","2021Q4") ~ "During Covid_19",
    Quarter %in% c("2022Q1", "2022Q2","2022Q3","2022Q4") ~ "Post-Covid 19"
  ))

simd %>% group_by( SIMD, Covid, AverageLengthOfStay) %>% 
  ggplot(aes(x = Covid, y = AverageLengthOfStay, fill = SIMD))+
  geom_boxplot()
```

