---
title: "Market Basket Analysis"
output: html_notebook
---

#Exercise

Load in some transaction data which has details on the items purchased in each 
transaction (where each transaction is uniquely identified by the InvoiceNo variable).

```{r}
library(tidyverse)
transactions <- read_csv("clean_data/online_retail_subset.csv")
```

We are interested in the purchase of two particular items:

item A

- ‘HEART OF WICKER SMALL’ (StockCode 22469)

item B

- ‘LARGE CAKE TOWEL PINK SPOTS’ (StockCode 21110)

1. Calculate the support for item A

```{r}
total_transactions <- transactions %>%
  summarise(n = n_distinct(InvoiceNo)) %>%
  flatten_dbl()

support_a <- transactions %>%
  filter(StockCode == 22469) %>%
  summarise(prop_invoices_with_item = n_distinct(InvoiceNo) / total_transactions)

support_a
```
Item A is purchased in 7.7% of transactions. 


2. Calculate the support and confidence for rule (A→B)

Going to join two tibbles containing the unique InvoiceNos of all the 
transactions involving item A, and the same for item B. Then join the tibbles 
on InvoiceNo to look for values in common: these will be transactions involving 
both items.

```{r}
trans_feat_a <- transactions %>%
  filter(StockCode == "22469") %>%
  distinct(InvoiceNo)

trans_feat_b <- transactions %>%
  filter(StockCode == "21110") %>%
  distinct(InvoiceNo)

trans_feat_a_b <- trans_feat_a %>%
  inner_join(trans_feat_b, by = "InvoiceNo")

support_a_b <- trans_feat_a_b %>%
  summarise(prop_with_a_b = n() / total_transactions)
support_a_b
```

```{r}
confidence_a_b <- support_a_b / support_a
confidence_a_b
```
Can find this out in a single pipeline

```{r}
support_a_b <- transactions %>%
  filter(StockCode == "22469" | StockCode == "21110") %>%
  group_by(InvoiceNo, StockCode) %>% #group by in case cases where more than one stock in a purchase
  summarise(count_of_item = n()) %>% 
  group_by(InvoiceNo) %>%
  summarise(count_of_A_and_B = n()) %>%
  filter(count_of_A_and_B > 1) %>%
  summarise(prop_with_A_and_B = n()/total_transactions)
```
```{r}
support_a_b 
```
```{r}
confidence_a_b <- support_a_b / support_a

confidence_a_b
```
If someone buys item A there is an 4.6% probability that they will then buy item B. 

3. Calculate the lift for (A→B)

```{r}
support_b <- transactions %>%
  filter(StockCode == 21110) %>%
  summarise(prop_invoices_with_item = n_distinct(InvoiceNo)/total_transactions)

lift_a_b <- support_a_b / (support_a * support_b)

lift_a_b
```
The lift is high, so A and B are more likely to be sold together, but from the 
support of A (7.7%) and confidence of (A→B) (4.6%), we see that happens at quite
a low frequency.




